---
title: "DM_Modelling"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Load packages
```{r}
library(pacman)
pacman::p_load(tidyverse, ggplot2, brms, boot, tidyverse, bayesplot)
```

# Load data
```{r}
# Set working directory
setwd("~/Documents/University/MA_COGNITIVESCIENCE/1SEMESTER/Decision Making/Exam_Analysis")
# Load data
d = read_csv("MatchingPennies_Processed.csv")
```

# Strategy Data: Getting subsets for data based on conditions: Double Alternation/Random Bias, SZ/HC
```{r}
# Double Alternation - SZ
DA_SZ = d %>% filter(subject_group == "SZ" & bot_strategy == "DoubleAlternation") %>%
  na.omit(subject_left)
  
# Double Alternation - HC
DA_HC = d %>% filter(subject_group == "HC" & bot_strategy == "DoubleAlternation") %>%
  na.omit(subject_left)

# Random Bias - SZ
RB_SZ = d %>% filter(subject_group == "SZ" & bot_strategy == "SimpleReversal") %>%
  na.omit(subject_left)

# Random Bias - SZ
RB_HC = d %>% filter(subject_group == "HC" & bot_strategy == "SimpleReversal") %>%
  na.omit(subject_left)
```

# Prior Check Function
```{r}
# Function for prior checks
prior_checking <- function(formula, data, prior, family, name){

  # fit priors
  prior_model <- brm(
    formula=formula,
    data=data,
    family = family,
    prior=prior,
    sample_prior = "only",
    chains = 4,
    cores = 3,
    seed = 2,
    backend = "cmdstanr",
    threads = threading(2),
    file = paste0("prior_models/", name, "_model.rds"),
    file_refit = "on_change",
    control = list(adapt_delta = 0.99,
                   max_treedepth = 20))
  
  # pp check, looks weird because only 0 and 1 are possible
  prior_checks = pp_check(prior_model)
  png(paste0("prior_models/", name, "_prior_plot.png"))
  print(prior_checks)
  dev.off()
  
  # generate predictions, to plot distribution of actual outcome values, centered at 50%,
  # average expectations, expectations only in terms of the average
  data$predictions <- predict(prior_model)[, 1]
  prior_predictions = ggplot(data, aes(predictions)) +
    theme_classic() + geom_density(alpha = 0.4)
  ggplot2::ggsave(paste0("prior_models/", name, "_predictions_plot.png"), prior_predictions)
  
  # average expectations, including the uncertainty
  png(paste0("prior_models/", name, "_predictions_noise_plot.png"))
  print(hist(inv_logit_scaled(posterior_linpred(prior_model, summary = FALSE))))
  dev.off()
}
``` 

# Posterior Check Function
```{r}
# Function for posterior update plots
plot_posterior_updates <- function(prior_name, posterior_name, posterior_samples, name){
    title = paste("Distributions of", prior_name, "(light) and", posterior_name, "(dark)")
    filename = paste0("posterior_models/", name, "_posterior_updates_", prior_name, posterior_name, ".png")
    plot = ggplot(posterior_samples) +
      theme_classic() +
      geom_density(aes_string(prior_name), fill = "lightsteelblue", alpha = 0.6) +
      geom_density(aes_string(posterior_name), fill = "royalblue4", alpha = 0.8) +
      labs(x="", y="", color="Legend", title=title)
    ggsave(filename, plot)
    print("saved")
}

# Function for posterior checks
posterior_checking <- function(formula, data, prior, family, name){
  # fit model
  posterior_model <- brm(
    formula=formula,
    data=data,
    family = family,
    prior=prior,
    sample_prior = TRUE,
    chains = 4,
    cores = 3,
    seed = 2,
    backend = "cmdstanr",
    threads = threading(2),
    file = paste0("posterior_models/", name, "_model.rds"),
    file_refit = "on_change",
    control = list(adapt_delta = 0.99,
                   max_treedepth = 20))
  
  # save model summary
  sink(paste0("posterior_models/", name, "_summary.txt"))
  print(summary(posterior_model))
  sink()
  
  # pp check, looks weird because only 0 and 1 are possible
  posterior_checks = pp_check(posterior_model)
  png(paste0("posterior_models/", name, "_posterior_plot.png"))
  print(posterior_checks)
  dev.off()
  
  # generate predictions, to plot distribution of actual outcome values, centered at 50%,
  # average expectations, expectations only in terms of the average
  data$predictions <- predict(posterior_model)[, 1]
  prior_predictions = ggplot(data, aes(predictions)) +
    theme_classic() + geom_density(alpha = 0.4)
  ggplot2::ggsave(paste0("posterior_models/", name, "_predictions_plot.png"), prior_predictions)
  
  # average expectations, including the uncertainty
  png(paste0("posterior_models/", name, "_predictions_noise_plot.png"))
  print(hist(inv_logit_scaled(posterior_linpred(posterior_model, summary = FALSE))))
  dev.off()
  
  # get posterior samples for plot 
  posterior_samples <- posterior_samples(posterior_model)
  plot_posterior_updates("prior_Intercept", "b_Intercept", posterior_samples, name)
  plot_posterior_updates("prior_sd_subject", "sd_subject__Intercept", posterior_samples, name)

  if ("b_left_prob" %in% variables(posterior_model)){
    plot_posterior_updates("prior_b", "b_left_prob", posterior_samples, name)
    plot_posterior_updates("prior_sd_subject", "sd_subject__left_prob", posterior_samples, name)
    plot_posterior_updates("prior_cor_subject", "cor_subject__Intercept__left_prob", posterior_samples, name)}
  
  if ("b_da_left_prob" %in% variables(posterior_model)){
    plot_posterior_updates("prior_b", "b_da_left_prob", posterior_samples, name)
    plot_posterior_updates("prior_sd_subject", "sd_subject__da_left_prob", posterior_samples, name)
    plot_posterior_updates("prior_cor_subject", "cor_subject__Intercept__da_left_prob", posterior_samples, name)}
  
  if ("b_stay_bias" %in% variables(posterior_model)){
    plot_posterior_updates("prior_b", "b_stay_bias", posterior_samples, name)
    plot_posterior_updates("prior_b", "b_leave_bias", posterior_samples, name)
    plot_posterior_updates("prior_sd_subject", "sd_subject__stay_bias", posterior_samples, name)
    plot_posterior_updates("prior_sd_subject", "sd_subject__leave_bias", posterior_samples, name)
    plot_posterior_updates("prior_cor_subject", "cor_subject__Intercept__stay_bias", posterior_samples, name)
    plot_posterior_updates("prior_cor_subject", "cor_subject__Intercept__leave_bias", posterior_samples, name)
    plot_posterior_updates("prior_cor_subject", "cor_subject__stay_bias__leave_bias", posterior_samples, name)
  }
}

```

# Defining the learning strategies 
```{r}
randombias_f <- bf(
  subject_left ~ 1 + (1 | subject))

biasednash_f <- bf(
  subject_left ~ 1 + left_prob + (1 + left_prob | subject))

doublealternation_f <- bf(
  subject_left ~ 1 + da_left_prob + (1 + da_left_prob | subject))

winstay_f <- bf(
  subject_left ~ 1 + stay_bias + leave_bias + (1 + stay_bias + leave_bias | subject))
```

# Simulate Priors
```{r}
# intercept
hist(inv_logit_scaled(rnorm(10000,0,1.5)))
# beta
hist(inv_logit_scaled(rnorm(10000,0,1.5) + rnorm(10000,0,1)))
```

# Simple Bias Model Prior and Posterior Checks
```{r}
# getting priors, for sample data set
get_prior(randombias_f, DA_SZ, family = bernoulli) 

# defining priors
randombias_prior <- c(
  prior(normal(0, 1.5), class = Intercept),
  prior(normal(0, 0.5), class = sd))

# run prior checks 
prior_checking(randombias_f, DA_SZ, randombias_prior, bernoulli, "DA_SZ/randombias_")
prior_checking(randombias_f, DA_HC, randombias_prior, bernoulli, "DA_HC/randombias_")
prior_checking(randombias_f, RB_SZ, randombias_prior, bernoulli, "RB_SZ/randombias_")
prior_checking(randombias_f, RB_HC, randombias_prior, bernoulli, "RB_HC/randombias_")

# run posterior checks
posterior_checking(randombias_f, DA_SZ, randombias_prior, bernoulli, "DA_SZ/randombias_")
posterior_checking(randombias_f, DA_HC, randombias_prior, bernoulli, "DA_HC/randombias_")
posterior_checking(randombias_f, RB_SZ, randombias_prior, bernoulli, "RB_SZ/randombias_")
posterior_checking(randombias_f, RB_HC, randombias_prior, bernoulli, "RB_HC/randombias_")
```

# Biased Nash Model Prior and Posterior Checks
```{r}
# getting priors, for sample data set
get_prior(biasednash_f, DA_SZ, family = bernoulli)

# defining priors
biasednash_prior <- c(
  prior(normal(0, 1.5), class = Intercept),
  prior(normal(0, 1.5), class = b),
  prior(normal(0, 0.5), class = sd),
  prior(lkj(2), class = cor))

# run prior checks 
prior_checking(biasednash_f, DA_SZ, biasednash_prior, bernoulli, "DA_SZ/biasednash_")
prior_checking(biasednash_f, DA_HC, biasednash_prior, bernoulli, "DA_HC/biasednash_")
prior_checking(biasednash_f, RB_SZ, biasednash_prior, bernoulli, "RB_SZ/biasednash_")
prior_checking(biasednash_f, RB_HC, biasednash_prior, bernoulli, "RB_HC/biasednash_")

# run posterior checks
posterior_checking(biasednash_f, DA_SZ, biasednash_prior, bernoulli, "DA_SZ/biasednash_")
posterior_checking(biasednash_f, DA_HC, biasednash_prior, bernoulli, "DA_HC/biasednash_")
posterior_checking(biasednash_f, RB_SZ, biasednash_prior, bernoulli, "RB_SZ/biasednash_")
posterior_checking(biasednash_f, RB_HC, biasednash_prior, bernoulli, "RB_HC/biasednash_")
```

# Double Alternation Model Prior and Posterior Checks
```{r}
# getting priors, for sample data set
get_prior(doublealternation_f, DA_SZ, family = bernoulli)

# defining priors
doublealternation_prior <- c(
  prior(normal(0, 1.5), class = Intercept),
  prior(normal(0, 1.5), class = b),
  prior(normal(0, 0.5), class = sd),
  prior(lkj(2), class = cor))

# run prior checks 
prior_checking(doublealternation_f, DA_SZ, doublealternation_prior, bernoulli, "DA_SZ/doublealternation_")
prior_checking(doublealternation_f, DA_HC, doublealternation_prior, bernoulli, "DA_HC/doublealternation_")
prior_checking(doublealternation_f, RB_SZ, doublealternation_prior, bernoulli, "RB_SZ/doublealternation_")
prior_checking(doublealternation_f, RB_HC, doublealternation_prior, bernoulli, "RB_HC/doublealternation_")

# run posterior checks
posterior_checking(doublealternation_f, DA_SZ, doublealternation_prior, bernoulli, "DA_SZ/doublealternation_")
posterior_checking(doublealternation_f, DA_HC, doublealternation_prior, bernoulli, "DA_HC/doublealternation_")
posterior_checking(doublealternation_f, RB_SZ, doublealternation_prior, bernoulli, "RB_SZ/doublealternation_")
posterior_checking(doublealternation_f, RB_HC, doublealternation_prior, bernoulli, "RB_HC/doublealternation_")
```

# Win Stay Model Prior and Posterior Checks
```{r}
winstay_f <- bf(
  subject_left ~ 1 + stay_bias + leave_bias + (1 + stay_bias + leave_bias | subject))

# getting priors, for sample data set
get_prior(winstay_f, DA_SZ, family = bernoulli)

# defining priors
winstay_prior <- c(
  prior(normal(0, 1), class = Intercept),
  prior(normal(0, 1), class = b),
  prior(normal(0, 0.5), class = sd),
  prior(lkj(2), class = cor))

# run prior checks 
prior_checking(winstay_f, DA_SZ, winstay_prior, bernoulli, "DA_SZ/winstay_")
prior_checking(winstay_f, DA_HC, winstay_prior, bernoulli, "DA_HC/winstay_")
prior_checking(winstay_f, RB_SZ, winstay_prior, bernoulli, "RB_SZ/winstay_")
prior_checking(winstay_f, RB_HC, winstay_prior, bernoulli, "RB_HC/winstay_")

# run posterior checks
posterior_checking(winstay_f, DA_SZ, winstay_prior, bernoulli, "DA_SZ/winstay_")
posterior_checking(winstay_f, DA_HC, winstay_prior, bernoulli, "DA_HC/winstay_")
posterior_checking(winstay_f, RB_SZ, winstay_prior, bernoulli, "RB_SZ/winstay_")
posterior_checking(winstay_f, RB_HC, winstay_prior, bernoulli, "RB_HC/winstay_")
```


# Mixture Model
```{r}

mixture <- mixture(bernoulli, bernoulli, bernoulli, bernoulli, order = "none")
# 1: simple bias
# 2: biased nash
# 3: double alternation
# 4: win stay - loose shift

mixture_f <- bf(
  # base
  subject_left ~ 1,
  # strategies
  mu1 ~ 1 + (1 | subject),
  mu2 ~ 1 + left_prob + (1 + left_prob | subject),
  mu3 ~ 1 + da_left_prob + (1 + da_left_prob | subject),
  mu4 ~ 1 + stay_bias + leave_bias + (1 + stay_bias + leave_bias | subject))

# get priors
get_prior(mixture_f, DA_SZ, mixture)

# define priors
mixture_prior <- c(
  # intercepts
  prior(normal(0,1.5), class=Intercept, dpar=mu1),
  prior(normal(0,1.5), class=Intercept, dpar=mu2),
  prior(normal(0,1.5), class=Intercept, dpar=mu3),
  prior(normal(0,1), class=Intercept, dpar=mu4),
  # betas
  prior(normal(0,1.5), class=b, dpar=mu2),
  prior(normal(0,1.5), class=b, dpar=mu3),
  prior(normal(0,1), class=b, dpar=mu4),
  # sd's
  prior(normal(0,0.5), class = sd, dpar=mu1),
  prior(normal(0,0.5), class = sd, dpar=mu2),
  prior(normal(0,0.5), class = sd, dpar=mu3),
  prior(normal(0,0.5), class = sd, dpar=mu4),
  # correlations
  prior(lkj(2), class = cor)
)

```

# Mixture Model of DA_SZ
```{r}

data = RB_SZ
name = "RB_SZ"

# running prior mixture model
mixture_model_prior <- brm(
  mixture_f,
  data,
  family = mixture,
  mixture_prior,
  sample_prior = "only",
  chains = 4,
  cores = 3,
  seed=2,
  backend = "cmdstanr",
  threads = threading(2),
  file = paste0("prior_mixture_models/", name),
  file_refit = "on_change",
  control = list(
    adapt_delta = 0.99,
    max_treedepth = 20
  )
)

# running posterior mixture model
mixture_model_posterior <- brm(
  mixture_f,
  data,
  family = mixture,
  mixture_prior,
  sample_prior = T,
  chains = 4,
  cores = 3,
  seed=2,
  backend = "cmdstanr",
  threads = threading(2),
  file = paste0("posterior_mixture_models/", name),
  file_refit = "on_change",
  control = list(
    adapt_delta = 0.99,
    max_treedepth = 20
  )
)
```

# Prior Checks of Mixture Models
```{r}

# define dataset
name = "DA_SZ"
data = DA_SZ

prior_model = readRDS(paste0("prior_mixture_models/", name, ".rds"))

# simple ppcheck
prior_checks = pp_check(prior_model)
png(paste0("prior_mixture_models/", name, "_prior_plot.png"))
print(prior_checks)
dev.off()

# predictions
data$predictions <- predict(prior_model)[, 1]
prior_predictions = ggplot(data, aes(predictions)) +
  theme_classic() + geom_density(alpha = 0.4)
ggplot2::ggsave(paste0("prior_mixture_models/", name, "_predictions_plot.png"), prior_predictions)
  
# predictions and noise
png(paste0("prior_mixture_models/", name, "_predictions_noise_plot.png"))
print(hist(inv_logit_scaled(posterior_linpred(prior_model, summary = FALSE))))
dev.off()

```

# Posterior Checks of Mixture Models
```{r}

name = "DA_SZ"
data = DA_SZ

posterior_model = readRDS(paste0("posterior_mixture_models/", name, ".rds"))

# summary
sink(paste0("posterior_mixture_models/", name, "_summary.txt"))
print(summary(posterior_model))
sink()

# pp check, looks weird because only 0 and 1 are possible
posterior_checks = pp_check(posterior_model)
png(paste0("posterior_mixture_models/", name, "_posterior_plot.png"))
print(posterior_checks)
dev.off()

# generate predictions, to plot distribution of actual outcome values, centered at 50%,
# average expectations, expectations only in terms of the average
data$predictions <- predict(posterior_model)[, 1]
posterior_predictions = ggplot(data, aes(predictions)) +
  theme_classic() + geom_density(alpha = 0.4)
ggplot2::ggsave(paste0("posterior_mixture_models/", name, "_predictions_plot.png"), posterior_predictions)

# prior and posterior update plots
posterior_s = posterior_samples(posterior_model)

# theta1
title = paste("Distributions of prior_theta1 (light) and theta1 (dark)")
theta1 = ggplot(posterior_s) +
    theme_classic() +
    geom_density(aes(`prior_theta[1]`), fill = "lightsteelblue", alpha = 0.6) +
    geom_density(aes(`theta[1]`), fill = "royalblue4", alpha = 0.8) +
    labs(x="", y="", color="Legend", title=title)
ggplot2::ggsave(paste0("posterior_mixture_models/", name, "_update_prior_theta1.png"), theta1)

# theta2
title = paste("Distributions of prior_theta2 (light) and theta2 (dark)")
theta2 = ggplot(posterior_s) +
    theme_classic() +
    geom_density(aes(`prior_theta[2]`), fill = "lightsteelblue", alpha = 0.6) +
    geom_density(aes(`theta[2]`), fill = "royalblue4", alpha = 0.8) +
    labs(x="", y="", color="Legend", title=title)
ggplot2::ggsave(paste0("posterior_mixture_models/", name, "_update_prior_theta2.png"), theta2)

# theta3
title = paste("Distributions of prior_theta3 (light) and theta3 (dark)")
theta3 = ggplot(posterior_s) +
    theme_classic() +
    geom_density(aes(`prior_theta[3]`), fill = "lightsteelblue", alpha = 0.6) +
    geom_density(aes(`theta[3]`), fill = "royalblue4", alpha = 0.8) +
    labs(x="", y="", color="Legend", title=title)
ggplot2::ggsave(paste0("posterior_mixture_models/", name, "_update_prior_theta3.png"), theta3)

# theta4
title = paste("Distributions of prior_theta4 (light) and theta4 (dark)")
theta4 = ggplot(posterior_s) +
    theme_classic() +
    geom_density(aes(`prior_theta[4]`), fill = "lightsteelblue", alpha = 0.6) +
    geom_density(aes(`theta[4]`), fill = "royalblue4", alpha = 0.8) +
    labs(x="", y="", color="Legend", title=title)
ggplot2::ggsave(paste0("posterior_mixture_models/", name, "_update_prior_theta4.png"), theta4)

```

# Plots
```{r}

# load data
SZ_model = readRDS("posterior_mixture_models/RB_SZ.rds")
HC_model = readRDS("posterior_mixture_models/RB_HC.rds")

# get estimates of SZ
SZ_summary = as.data.frame(summary(SZ_model)$spec_pars)
SZ_summary <- cbind(theta = rownames(SZ_summary), SZ_summary)
rownames(SZ_summary) <- 1:nrow(SZ_summary)
SZ_summary$group = "SCZ"

# get estimates of HC
HC_summary = as.data.frame(summary(HC_model)$spec_pars)
HC_summary <- cbind(theta = rownames(HC_summary), HC_summary)
rownames(HC_summary) <- 1:nrow(HC_summary)
HC_summary$group = "HC"

# bind estimates together and rename columns
summaries = rbind(SZ_summary, HC_summary)
colnames(summaries) = c("theta", "estimate", "error", "lowerCI", "upperCI", "rhat", "bulkESS", "tailESS", "group")


# rename thetas
summaries$theta = ifelse(summaries$theta == "theta1", "RB",
                         ifelse(summaries$theta == "theta2", "BN",
                                ifelse(summaries$theta == "theta3", "DA",
                                       "WSLS")))

# relevel thetas
summaries$theta <- factor(summaries$theta, levels = c("RB", "BN", "DA", "WSLS"))

# rename groups
summaries$group = ifelse(summaries$group == "SCZ", "Schizophrenia Subjects (SCZ)", 
                         "Healthy Control Subjects (HC)")

summaries$group <- factor(summaries$group, levels = c("Schizophrenia Subjects (SCZ)", "Healthy Control Subjects (HC)"))

# make plot
ggplot(summaries, aes(x=theta, y=estimate, fill = theta)) + 
  geom_bar(stat="identity", position=position_dodge()) +
  geom_errorbar(aes(ymin=lowerCI, ymax=upperCI), width=.2,position=position_dodge(.9)) +
  labs(title = "Mixing Proportions of SCZ and HC against Random Bias Opponent", 
       x="Adaptation Strategy", y = "Mixing Proportions in %", legend=FALSE) +
  theme_bw() +
  scale_fill_manual(name = "", values=c("#fd7f6f","#7eb0d5", "#b2e061", "#ffb55a"), 
                    labels = c("Random Bias (RB)", "Biased Nash (BN)", "Double Alternation (DA)", "Win-Stay/Lose-Switch (WSLS)")) +
  theme(plot.title = element_text(hjust = 0.5), legend.position="bottom") +
  facet_wrap("~group")
```


